Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2020-07-25 article_springer/article_springer.Rmd
Add lim to Defintion of Regular varying functions
Merge Remark after proof into one
Addressing comment 4 of Referee 2 in the article

$$
\frac{f(\lambda x)}{f(x)} = \lambda^\rho \quad \text{ for all } \lambda > 0.
$$

@@

:[@MeerschaertStoev08] proofed a limit theorem for the maxima of iid random variables separated by infinite mean waiting times. They showed that the duration time of the limit process is Mittag-Leffler distributed. [@Basrak2014] described the asymptotic distribution under similar assumptions of all upper orders statistics using point processes.

@@

: If $\beta = 1$, the result of the Theorem above is standard, see e.g. 
Equation (2.2) in @Gut1999. In @Anderson1987  a similar result is shown with a different choice of scaling constant.

@@

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple] based on the $r+1$ upper order statistics
\begin{align}\label{eq:Hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align}
where $X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}$ denote the order statistics in decreasing order of a sequence $X_1,\ldots,X_n$. The Hill estimator only performs reliably well for distributions close to Pareto [see @Resnick97]. It heavily depends on the choice of $r$ in \eqref{eq:Hill}, and hence on the number of order statistics entering \eqref{eq:Hill}.  Figure \ref{fig:Hillplots} shows 100 Hill plots (grey thin lines) for simulated Mittag-Leffler data with tail parameter $\beta=0.8$ and scale $\sigma=1$, for a sample size of $n=200$ (left) and $n=1000$ (right). The dark lines are the means and the red dotted lines show the true tail
parameter. To deduce the correct tail parameter estimates from one of those Hill plots is virtually impossible. To estimate the tail parameter of a Mittag-Leffler distribution with the Hill estimator, the sample size has to be much larger and one has to choose a small tuning parameter $r$. In case that only events above a high threshold are recorded, $n=10000$ events are extremely unrealistic. There is a vast literature on bias-reduced versions of the Hill estimator and on the optimal choice of $r$, mainly based on imposed second-order regular variation conditions [see e.g. @beirlantBook, Ch.4.4 and 4.7 and references in it]. Also, there are hundreds other tail-index estimators not considered in this article, since this would go beyond the scope of our article. For example the moment estimator [@dekkers1989moment], the Pickands estimator [@pickands1975statistical], the QQ-estimator [@Kratz96], to name a few. A nice overview can be found in [@fedotenkov2018review]. However, we know from the Theorem in Section \ref{sec:scaling} that the inter-exceedance times are asymptotically Mittag-Leffler distributed, and hence it is unsurprising that general tail (and scale) estimation methods that do not take into account this structure are statistically less efficient [@kozubowski2001]. Finally, the Hill estimator completely fails in the important special case where the inter-arrival times are exponentially distributed and not heavy-tailed. For these reasons, we propose to use the log-Moment estimator or the MLE for Mittag-Leffler distributions.

@@

The performance of log-Moment and Maximum-Likelihood estimators for the tail resp. the scale parameter is shown in Figure \ref{Fig:TailSimu} resp. \ref{Fig:ScaleSimu}. Both estimators show good performance, with a slight advantage for the Maximum-Likelihood estimator. This advantage is paid for with a higher computational cost. Moreover, the bottom row in Figure \ref{Fig:TailSimu} shows clearly that the log-Moment and Maximum-Likelihood estimators generalize to the exponential case ($\beta = 1$) whereas the Hill-estimator would not.

>>>>

$$
\underset{x\rightarrow \infty}{lim}\frac{f(\lambda x)}{f(x)} = \lambda^\rho \quad \text{ for all } \lambda > 0.
$$

@@

@@

: If $\beta = 1$, the result of the Theorem above is standard, see e.g. 
Equation (2.2) in @Gut1999. In @Anderson1987  a similar result is shown with a different choice of scaling constant. [@MeerschaertStoev08] proofed a limit theorem for the maxima of iid random variables separated by infinite mean waiting times. They showed that the duration time of the limit process is Mittag-Leffler distributed. [@Basrak2014] described the asymptotic distribution under similar assumptions of all upper orders statistics using point processes.

@@

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple] based on the $r+1$ upper order statistics
\begin{align}\label{eq:Hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align}
where $X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}$ denote the order statistics in decreasing order of a sequence $X_1,\ldots,X_n$. The Hill estimator only performs reliably well for distributions close to Pareto [see @Resnick97]. It heavily depends on the choice of $r$ in \eqref{eq:Hill}, and hence on the number of order statistics entering \eqref{eq:Hill}.  Figure \ref{fig:Hillplots} shows 100 Hill plots (grey thin lines) for simulated Mittag-Leffler data with tail parameter $\beta=0.8$ and scale $\sigma=1$, for a sample size of $n=200$ (left) and $n=1000$ (right). The dark lines are the means and the red dotted lines show the true tail
parameter. To deduce the correct tail parameter estimates from one of those Hill plots is virtually impossible. To estimate the tail parameter of a Mittag-Leffler distribution with the Hill estimator, the sample size has to be much larger and one has to choose a small tuning parameter $r$. In case that only events above a high threshold are recorded, $n=10000$ events are extremely unrealistic. There is a vast literature on bias-reduced versions of the Hill estimator and on the optimal choice of $r$, mainly based on imposed second-order regular variation conditions [see e.g. @beirlantBook, Ch.4.4 and 4.7 and references in it]. Also, there are hundreds other tail-index estimators not considered in this article, since this would go beyond the scope of our article. For example the moment estimator [@dekkers1989moment], the Pickands estimator [@pickands1975statistical], the QQ-estimator [@Kratz96], to name a few. A nice overview can be found in [@fedotenkov2018review]. However, we know from the Theorem in Section \ref{sec:scaling} that the inter-exceedance times are asymptotically Mittag-Leffler distributed, and hence it is unsurprising that general tail (and scale) estimation methods that do not take into account this structure are statistically less efficient [@kozubowski2001] on the *thresholded* data. Finally, the Hill estimator completely fails in the important special case where the inter-arrival times are exponentially distributed and not heavy-tailed. For these reasons, we propose to use the log-Moment estimator or the MLE for Mittag-Leffler distributions.

@@

The performance of log-Moment and Maximum-Likelihood estimators for the tail resp. the scale parameter is shown in Figure \ref{Fig:TailSimu} resp. \ref{Fig:ScaleSimu}. Both estimators show good performance, with a slight advantage for the Maximum-Likelihood estimator. This advantage is paid for with a higher computational cost. Moreover, the bottom row in Figure \ref{Fig:TailSimu} shows clearly that the log-Moment and Maximum-Likelihood estimators generalize to the exponential case ($\beta = 1$) whereas the Hill-estimator would not.  
Notice that any data below a certain threshold gets discarded in our approach and hence need not to satisfy the iid assumption. In our simulation study we only simulated data from iid sequences of $(W_k)_{k \in \mathbb{N}}$, in real data situations its likely that there are dependencies or more complex data generating processes vanishing for higher thresholds. We only need the exceedances as well as the inter-exceedance times to be iid.
